---
title: "Lab 3"
author: "Anne-Marie Parkinson"
date: "January 23, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      message=F,
                      warnings=F)

# cant put message and warnings = F in the r setup area bc those are instructions for only this r chunk
```

# Load Packages

```{r}
library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
```


## Binary Logistic Regression 

add data
```{r}

gentoo <-  read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

```


join the data sets and tidy the data. They have the same info, but the species is different 
```{r}
penguins <- full_join(chinstrap, gentoo) %>% 
  clean_names() %>% 
  mutate (sex = str_to_lower(sex)) %>% 
  filter(sex %in% c("male", "female"))
  

# in the brackets for full join, just provide the name of the data sets
```


data exploration graph 
```{r}
ggplot(penguins, aes(x=sex, y=body_mass_g)) +
  geom_jitter(aes(color = sex, pch = sex)) +
  facet_wrap(~species)

```

manually assign the penguins as 1 or 0

chinstrap = 1
gentoo = 0

```{r}

penguin_bin <- penguins %>% 
  mutate (sp_binary = case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
    species == "Gentoo penguin (Pygoscelis papua)" ~ 0
  ))

# can also use str_detect or contains and only write part of the species (ex/ just chin or gentoo) 

```

when using case_when, if you want to assign remaining groups to one group (w/o having to type it out) can use this code at the end:

penguin_bin <- penguins %>% 
  mutate (sp_binary = case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
    TRUE ~ 2
  ))

class has to be the same as the other observations, cannot have factor and numeric in same column

## binomial logistic regression
```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g, 
                   data=penguin_bin, 
                   family = "binomial")

summary(penguin_blr)


```

# lets make some predictions 

1. what is the probability that a penguin is a chinstrap, if it weighs 4500g and is male?

```{r}
# create data frame 
df_m4500 <- data.frame(sex = "male",
                       body_mass_g = 4500)

#inputs have to be exactly the same as the info in the original data

# find log odds of this 4500g male being a chinstrap penguin

m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link")
m4500_logodds # not that intuitive, so lets use type=repsonse

m4500_probs <- predict(penguin_blr, newdata=df_m4500, type = "response")
m4500_probs # easy to understand answer


```

2. whats the probability that a penguin is a chinstrap if it is a 4000g female 

```{r}

df_f4000 <- data.frame(sex = "female",
                       body_mass_g = 4000)

# find log odds of this 4000g female being a chinstrap penguin
f4000_probs <- predict(penguin_blr, newdata=df_f4000, type = "response")
f4000_probs

```

lets make an entire data frmae and use it to make then visualie logistic regression outcomes
```{r}
# create mock data 
penguins_mock <- data.frame(
  body_mass_g = rep(seq(3000, 6000, length = 200), 2),
  sex = c(rep("male", 200), rep("female", 200))
)

#determine predictions
full_predict <- predict(penguin_blr, newdata= penguins_mock, type="response", se.fit=T)

# combine mock data and prediction values

final_df <- data.frame(penguins_mock, 
                       full_predict$fit, 
                       full_predict$se.fit)

colnames(final_df) <- c("penguin_mass", "sex", "probability", "se")
final_df

# plot 
ggplot(data=final_df, aes(x=penguin_mass, y=probability)) +
  geom_line(aes(color=sex)) +
  geom_ribbon(aes(ymin = probability - se, 
                  ymax = probability + se,
                  fill = sex),
              alpha = 0.3) +
  labs (title = "Probability of being a chinstrap")

```

## using spatial data

```{r}
# add data
cougars <- read_sf(dsn = here("cougar_connections"), layer= "ds1014") %>% 
  clean_names()

#sub data
large_corridors <- cougars %>% 
  filter (area_ac > 20000) %>% 
  select(area_ac)

plot(large_corridors)

# sub data
cougars_sub <- cougars %>% 
  select(elev_mean)

ggplot(cougars_sub) +
  geom_sf(aes(fill = elev_mean), color=NA)

tmap_mode("view")  # set the mode: theres 2 options, view (which is interactive), and plot (not interactive)

tm_shape(cougars_sub) +
  tm_fill("elev_mean")

# click on the stacked squares in the map to select a basemap
```




















